quollr: An R Package for Visualizing \(2-\text{D}\) Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-08 paper-quollr.tex
Added layouts for example
Added updated version

% !TeX root = RJwrapper.tex
\title{quollr: An R Package for Visualizing \(2-D\) Models from Non-linear Dimension Reductions in High Dimensional Space}

@@

\abstract{%
Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (p\text{-}D) by applying a non-linear transformation. However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper)parameter choices. It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures. The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper)parameter choices provide the most accurate representation of high-dimensional data. The \texttt{triangular\_3d\_data} data from the \CRANpkg{cardinalR} package is used to illustrate the algorithm and its application within the package.
}

@@

@@

@@

\subsection{Usage}\label{usage}

The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. We also want to load the built-in data sets \texttt{s\_curve\_noise\_training} and \texttt{s\_curve\_noise\_umap}.

\texttt{s\_curve\_noise\_training} is a \(3\text{-}D\) S-curve data set with additional four noise dimensions which is used to train the model. \texttt{s\_curve\_noise\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{s\_curve\_noise\_training} data set. Each data set contains a unique ID column.

\subsubsection{Scaling the data}\label{scaling-the-data}

The algorithm begins by scaling NLDR data to a standard scale using the \texttt{gen\_scaled\_data()} function. This function standardizes the data and provides the original limits of embeddings and the aspect ratio.

\begin{verbatim}
scaled_nldr_obj_scurve <- gen_scaled_data(data = s_curve_noise_umap)

s_curve_noise_umap_scaled <- scaled_nldr_obj_scurve$scaled_nldr

lim1 <- scaled_nldr_obj_scurve$lim1
lim2 <- scaled_nldr_obj_scurve$lim2
r2 <- diff(lim2)/diff(lim1) 
\end{verbatim}

The mains steps for the algorithm can be executed by the main function \texttt{fit\_highd\_model()}, or can be run separately for more flexibility. When constructing the \(2\text{-}D\) model, the user can choose either to fit the \(2\text{-}D\) model with hexagonal bin centroids or bin means using \texttt{is\_bin\_centroid} argument.

If a user would like to perform steps of the algorithm themselves, additional user input will be needed for the function that perform each step. For example, if the user wishes to use already binning data, the \texttt{extract\_hexbin\_centroids()} function can be used directly.

The number of bins along the x-axis, the ratio of the ranges of the original embedding components, the buffer amount as a proportion of data, and if \texttt{is\_rm\_lwd\_hex\ =\ TRUE}, benchmark value to remove low density hexagons are parameters that will be determined within \texttt{fit\_highd\_model()}, if they are not provided. They are created as they are needed throughout the following example. The function \texttt{fit\_highd\_model()} provides the fitted model in \(2\text{-}D\), and \(p\text{-}D\).

\begin{verbatim}
fit_highd_model(
  highd_data = s_curve_noise_training,
  nldr_data = s_curve_noise_umap_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.1,
  is_bin_centroid = TRUE
  )
\end{verbatim}

\begin{verbatim}
#> $df_bin
#> # A tibble: 86 x 8
#>    hb_id       x1    x2    x3        x4        x5        x6        x7
#>    <int>    <dbl> <dbl> <dbl>     <dbl>     <dbl>     <dbl>     <dbl>
#>  1    15  0.968   0.198  1.21  0.000518  0.00161   0.00240  -0.000547
#>  2    16  0.730   0.142  1.67 -0.000689  0.000495  0.00168   0.000794
#>  3    17  0.323   0.158  1.94 -0.00265  -0.00523  -0.00942  -0.00115 
#>  4    18 -0.203   0.154  1.98 -0.00269  -0.00107  -0.0382   -0.00167 
#>  5    19 -0.628   0.114  1.78 -0.000131 -0.00323   0.0347   -0.00288 
#>  6    27  0.987   0.579  1.14 -0.00124  -0.00121   0.0199   -0.00363 
#>  7    28  0.847   0.537  1.51  0.000374 -0.00215  -0.00901  -0.000548
#>  8    29  0.469   0.476  1.87  0.00117  -0.000439 -0.0191    0.000251
#>  9    30 -0.00747 0.491  1.99 -0.000570  0.00114  -0.000815  0.000786
#> 10    31 -0.525   0.449  1.84  0.00141  -0.000575 -0.0112    0.000714
#> # i 76 more rows
#> 
#> $df_bin_centroids
#> # A tibble: 86 x 6
#>    hexID   c_x      c_y bin_counts std_counts drop_empty
#>    <int> <dbl>    <dbl>      <int>      <dbl> <lgl>     
#>  1    15 0.158 0.000959         43     0.518  FALSE     
#>  2    16 0.262 0.000959         35     0.422  FALSE     
#>  3    17 0.365 0.000959         21     0.253  FALSE     
#>  4    18 0.468 0.000959         11     0.133  FALSE     
#>  5    19 0.571 0.000959          8     0.0964 FALSE     
#>  6    27 0.107 0.0904           10     0.120  FALSE     
#>  7    28 0.210 0.0904           56     0.675  FALSE     
#>  8    29 0.313 0.0904           43     0.518  FALSE     
#>  9    30 0.416 0.0904           54     0.651  FALSE     
#> 10    31 0.520 0.0904           42     0.506  FALSE     
#> # i 76 more rows
\end{verbatim}

\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) model}{Constructing the 2\textbackslash text\{-\}D model}}\label{constructing-the-2text-d-model}

Constructing the \(2\text{-}D\) model mainly contains (i) binning data, (ii) obtaining bin centroids, and (iii) indicating neighbors by line segments connecting centroids.

@@

\begin{verbatim}
calc_bins_y(
  bin1 = 12, 
  r2 = r2,
  q = 0.1
  )
\end{verbatim}

\begin{verbatim}
#> $bin2
#> [1] 13
#> 
#> $a1
#> [1] 0.1032943
#> 
#> $a2
#> [1] 0.08945548
\end{verbatim}

@@

\begin{verbatim}
hex_binning(
  data = s_curve_noise_umap_scaled, 
  bin1 = 12, 
  r2 = r2,
  q = 0.1
  )
\end{verbatim}

\begin{verbatim}
#> $a1
#> [1] 0.1032943
#> 
#> $a2
#> [1] 0.08945548
#> 
#> $bins
#> [1] 12 13
#> 
#> $start_point
#> [1] -0.10000000 -0.08849688
#> 
#> $centroids
#> # A tibble: 156 x 3
#>    hexID      c_x     c_y
#>    <int>    <dbl>   <dbl>
#>  1     1 -0.1     -0.0885
#>  2     2  0.00329 -0.0885
#>  3     3  0.107   -0.0885
#>  4     4  0.210   -0.0885
#>  5     5  0.313   -0.0885
#>  6     6  0.416   -0.0885
#>  7     7  0.520   -0.0885
#>  8     8  0.623   -0.0885
#>  9     9  0.726   -0.0885
#> 10    10  0.830   -0.0885
#> # i 146 more rows
#> 
#> $hex_poly
#> # A tibble: 936 x 3
#>    hex_poly_id        x       y
#>          <int>    <dbl>   <dbl>
#>  1           1 -0.1     -0.0289
#>  2           1 -0.152   -0.0587
#>  3           1 -0.152   -0.118 
#>  4           1 -0.1     -0.148 
#>  5           1 -0.0484  -0.118 
#>  6           1 -0.0484  -0.0587
#>  7           2  0.00329 -0.0289
#>  8           2 -0.0484  -0.0587
#>  9           2 -0.0484  -0.118 
#> 10           2  0.00329 -0.148 
#> # i 926 more rows
#> 
#> $data_hb_id
#> # A tibble: 3,750 x 4
#>     emb1   emb2    ID hb_id
#>    <dbl>  <dbl> <int> <int>
#>  1 0.270 0.839      1   125
#>  2 0.788 0.466      2    82
#>  3 0.771 0.319      3    69
#>  4 0.306 0.542      5    88
#>  5 0.549 0.806      6   127
#>  6 0.166 0.259      7    52
#>  7 0.672 0.596      9   104
#>  8 0.735 0.354     10    69
#>  9 0.839 0.467     11    82
#> 10 0.240 0.0903    12    28
#> # i 3,740 more rows
#> 
#> $std_cts
#> # A tibble: 86 x 3
#>    hb_id     n std_counts
#>    <int> <int>      <dbl>
#>  1    15    43     0.518 
#>  2    16    35     0.422 
#>  3    17    21     0.253 
#>  4    18    11     0.133 
#>  5    19     8     0.0964
#>  6    27    10     0.120 
#>  7    28    56     0.675 
#>  8    29    43     0.518 
#>  9    30    54     0.651 
#> 10    31    42     0.506 
#> # i 76 more rows
#> 
#> $tot_bins
#> [1] 156
#> 
#> $non_bins
#> [1] 86
#> 
#> $pts_bins
#> # A tibble: 86 x 2
#>    hb_id pts_list  
#>    <int> <list>    
#>  1    15 <int [43]>
#>  2    16 <int [35]>
#>  3    17 <int [21]>
#>  4    18 <int [11]>
#>  5    19 <int [8]> 
#>  6    27 <int [10]>
#>  7    28 <int [56]>
#>  8    29 <int [43]>
#>  9    30 <int [54]>
#> 10    31 <int [42]>
#> # i 76 more rows
#> 
#> attr(,"class")
#> [1] "hex_bin_obj"
\end{verbatim}

\begin{verbatim}
find_low_dens_hex(
  df_bin_centroids_all = df_bin_centroids, 
  bin1 = 12, 
  df_bin_centroids_low = df_bin_centroids_low
  )
\end{verbatim}

@@

\begin{verbatim}
find_lg_benchmark(
  distance_edges = distance_df, 
  distance_col = "distance"
  )
\end{verbatim}

@@

The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each bin to represent bin centroids. This transformation is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data and their corresponding hexagonal bin IDs as inputs.

\begin{verbatim}
umap_data_with_hb_id <- hb_obj$data_hb_id

df_all <- bind_cols(
  s_curve_noise_training |> dplyr::select(-ID), 
  umap_data_with_hb_id
  )

df_bin <- avg_highd_data(
  data = df_all
  )
\end{verbatim}

@@

\begin{verbatim}
predict_emb(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
\end{verbatim}

@@

\begin{verbatim}
glance(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
\end{verbatim}

@@

\begin{verbatim}
augment(
  highd_data = s_curve_noise_training, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
\end{verbatim}

@@

\begin{verbatim}
vis_lg_mesh(
  distance_edges = distance_df, 
  benchmark_value = 0.75, 
  tr_coord_df = tr_from_to_df, 
  distance_col = "distance"
  )
\end{verbatim}

\begin{verbatim}
vis_rmlg_mesh(
  distance_edges = distance_df, 
  benchmark_value = 0.75, 
  tr_coord_df = tr_from_to_df, 
  distance_col = "distance"
  )
\end{verbatim}

@@

\begin{verbatim}
show_langevitour(
  df = df_all, 
  df_b = df_bin, 
  df_b_with_center_data = df_bin_centroids, 
  benchmark_value = 0.75, 
  distance = distance_df, 
  distance_col = "distance", 
  use_default_benchmark_val = FALSE, 
  col_start = "x"
  )
\end{verbatim}

>>>>

% !TeX root = RJwrapper.tex
\title{quollr: An R Package for Visualizing \(2-\text{D}\) Models from Non-linear Dimension Reductions in High Dimensional Space}

@@

\abstract{%
Non-linear dimension reduction (NLDR) methods provide a low-dimensional representation of high-dimensional data (p\text{-}D) by applying a non-linear transformation. However, the complexity of the transformations and data structures can create wildly different representations depending on the method and (hyper)parameter choices. It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures. The R package \CRANpkg{quollr} has been developed as a new visual tool to determine which method and which (hyper)parameter choices provide the most accurate representation of high-dimensional data. The \texttt{scurve} data from the package is used to illustrate the algorithm. scRNA-seq data for Limb muscles of mice are used to discuss the usability of the package.
}

@@

\subsection{Installation}\label{installation}

@@

\subsection{Web site}\label{web-site}

More documentation of the package can be found at the web site \url{https://jayanilakshika.github.io/quollr/}.

@@

\subsection{Data sets}\label{data-sets}

Add a table of data sets that quollr have

The following demonstration of the package's functionality assumes \texttt{quollr} has been loaded. We also want to load the built-in data sets \texttt{scurve} and \texttt{scurve\_umap}.

\texttt{scurve} is a \(7\text{-}D\) simulated dataset. It is constructed by simulating \(5000\) observations from \(\theta \sim U(-3\pi/2, 3\pi/2)\), \(X_1 = \sin(\theta)\), \(X_2 \sim U(0, 2)\) (adding thickness to the S), \(X_3 = \text{sign}(\theta) \times (\cos(\theta) - 1)\). The remaining variables \(X_4, X_5, X_6, X_7\) are all uniform error, with small variance. \texttt{scurve\_umap} is the UMAP \(2\text{-}D\) embedding for \texttt{scurve} data with \texttt{n\_neighbors} is \(46\) and \texttt{min\_dist} is \(0.9\). Each data set contains a unique ID column that maps \texttt{scurve} and \texttt{scurve\_umap}.

\subsection{Main function}\label{main-function}

The mains steps for the algorithm can be executed by the main function \texttt{fit\_highd\_model()}, or can be run separately for more flexibility.

This function requires several parameters: the high-dimensional data (\texttt{highd\_data}), the emdedding data (\texttt{nldr\_data}), the number of bins along the x-axis (\texttt{bin1}), the buffer amount as a proportion of data (\texttt{q}), and benchmark value to extract high density hexagons (\texttt{benchmark\_highdens}). The function returns an object that includes the hexagonal object, the fitted model in both \(2\text{-}D\), and \(p\text{-}D\), and triangular mesh.

\begin{verbatim}
fit_highd_model(
  highd_data = scurve, 
  nldr_data = scurve_umap, 
  bin1 = 15, 
  q = 0.1, 
  benchmark_highdens = 5)
\end{verbatim}

\begin{verbatim}
#> $hb_obj
#> $a1
#> [1] 0.08326136
#> 
#> $a2
#> [1] 0.07210645
#> 
#> $bins
#> [1] 15 16
#> 
#> $start_point
#> [1] -0.10000000 -0.08923607
#> 
#> $centroids
#> # A tibble: 240 x 3
#>    hexID     c_x     c_y
#>    <int>   <dbl>   <dbl>
#>  1     1 -0.1    -0.0892
#>  2     2 -0.0167 -0.0892
#>  3     3  0.0665 -0.0892
#>  4     4  0.150  -0.0892
#>  5     5  0.233  -0.0892
#>  6     6  0.316  -0.0892
#>  7     7  0.400  -0.0892
#>  8     8  0.483  -0.0892
#>  9     9  0.566  -0.0892
#> 10    10  0.649  -0.0892
#> # i 230 more rows
#> 
#> $hex_poly
#> # A tibble: 1,440 x 3
#>    hex_poly_id       x       y
#>          <int>   <dbl>   <dbl>
#>  1           1 -0.1    -0.0412
#>  2           1 -0.142  -0.0652
#>  3           1 -0.142  -0.113 
#>  4           1 -0.1    -0.137 
#>  5           1 -0.0584 -0.113 
#>  6           1 -0.0584 -0.0652
#>  7           2 -0.0167 -0.0412
#>  8           2 -0.0584 -0.0652
#>  9           2 -0.0584 -0.113 
#> 10           2 -0.0167 -0.137 
#> # i 1,430 more rows
#> 
#> $data_hb_id
#> # A tibble: 5,000 x 4
#>     emb1  emb2    ID hexID
#>    <dbl> <dbl> <int> <int>
#>  1 0.707 0.839     1   205
#>  2 0.231 0.401     2   109
#>  3 0.232 0.215     3    65
#>  4 0.790 0.564     4   146
#>  5 0.761 0.551     5   146
#>  6 0.445 0.721     6   172
#>  7 0.900 0.137     7    58
#>  8 0.247 0.392     8   110
#>  9 0.325 0.542     9   141
#> 10 0.278 0.231    10    80
#> # i 4,990 more rows
#> 
#> $std_cts
#> # A tibble: 142 x 3
#>    hexID bin_counts std_counts
#>    <int>      <int>      <dbl>
#>  1    21         12     0.16  
#>  2    22         17     0.227 
#>  3    23         22     0.293 
#>  4    24         10     0.133 
#>  5    25          7     0.0933
#>  6    26         12     0.16  
#>  7    27          1     0.0133
#>  8    36         42     0.56  
#>  9    37         42     0.56  
#> 10    38         44     0.587 
#> # i 132 more rows
#> 
#> $tot_bins
#> [1] 240
#> 
#> $non_bins
#> [1] 142
#> 
#> $pts_bins
#> # A tibble: 142 x 2
#>    hexID pts_list  
#>    <int> <list>    
#>  1    21 <int [12]>
#>  2    22 <int [17]>
#>  3    23 <int [22]>
#>  4    24 <int [10]>
#>  5    25 <int [7]> 
#>  6    26 <int [12]>
#>  7    27 <int [1]> 
#>  8    36 <int [42]>
#>  9    37 <int [42]>
#> 10    38 <int [44]>
#> # i 132 more rows
#> 
#> attr(,"class")
#> [1] "hex_bin_obj"
#> 
#> $model_highd
#> # A tibble: 130 x 8
#>    hexID      x1    x2    x3         x4        x5       x6        x7
#>    <int>   <dbl> <dbl> <dbl>      <dbl>     <dbl>    <dbl>     <dbl>
#>  1    21 -0.992   1.91 1.11  -0.000427   0.000624  0.00749  0.00105 
#>  2    22 -0.906   1.93 1.41  -0.0000183  0.00331  -0.0204  -0.000363
#>  3    23 -0.680   1.93 1.72  -0.000810  -0.00259  -0.00449  0.00153 
#>  4    24 -0.272   1.93 1.96   0.00251    0.00668  -0.0460   0.00128 
#>  5    25  0.0760  1.93 2.00   0.00876    0.00447   0.00851 -0.00195 
#>  6    26  0.461   1.93 1.89  -0.00478    0.00492   0.00835  0.00172 
#>  7    36 -0.985   1.75 0.853 -0.00202    0.000397  0.00331  0.000338
#>  8    37 -0.980   1.66 1.17  -0.000374  -0.00154   0.0165   0.000126
#>  9    38 -0.821   1.64 1.56  -0.000459   0.000538 -0.0123   0.000780
#> 10    39 -0.484   1.68 1.87   0.00313    0.00241   0.00823 -0.00117 
#> # i 120 more rows
#> 
#> $model_2d
#> # A tibble: 130 x 5
#>    hexID   c_x     c_y bin_counts std_counts
#>    <int> <dbl>   <dbl>      <dbl>      <dbl>
#>  1    21 0.358 -0.0171         12     0.16  
#>  2    22 0.441 -0.0171         17     0.227 
#>  3    23 0.524 -0.0171         22     0.293 
#>  4    24 0.608 -0.0171         10     0.133 
#>  5    25 0.691 -0.0171          7     0.0933
#>  6    26 0.774 -0.0171         12     0.16  
#>  7    36 0.316  0.0550         42     0.56  
#>  8    37 0.400  0.0550         42     0.56  
#>  9    38 0.483  0.0550         44     0.587 
#> 10    39 0.566  0.0550         39     0.52  
#> # i 120 more rows
#> 
#> $trimesh_data
#> # A tibble: 324 x 6
#>     from    to x_from  y_from  x_to   y_to
#>    <int> <int>  <dbl>   <dbl> <dbl>  <dbl>
#>  1    16    26 0.275   0.127  0.233 0.199 
#>  2    25    26 0.150   0.199  0.233 0.199 
#>  3    25    37 0.150   0.199  0.191 0.271 
#>  4    36    49 0.108   0.271  0.150 0.343 
#>  5    36    37 0.108   0.271  0.191 0.271 
#>  6     1     7 0.358  -0.0171 0.316 0.0550
#>  7    48    49 0.0665  0.343  0.150 0.343 
#>  8    37    38 0.191   0.271  0.275 0.271 
#>  9    26    27 0.233   0.199  0.316 0.199 
#> 10    27    38 0.316   0.199  0.275 0.271 
#> # i 314 more rows
\end{verbatim}

\subsection{\texorpdfstring{Constructing the \(2\text{-}D\) Model}{Constructing the 2\textbackslash text\{-\}D Model}}\label{constructing-the-2text-d-model}

Constructing the \(2\text{-}D\) model primarily involves (i) scaling the NLDR data, (ii) binning the data, (iii) obtaining bin centroids, (iv) connecting centroids with line segments to indicate neighbors, and (v) Remove low-density hexagons.

\subsubsection{Scaling the Data}\label{scaling-the-data}

The algorithm starts by scaling the NLDR data to a standard range using the \texttt{gen\_scaled\_data()} function. This function standardizes the data so that the first embedding ranges from \(0\) to \(1\), while the second embedding scales from \(0\) to the maximum value of the second embedding. The output includes the scaled NLDR data along with the original limits of the embeddings.

\begin{verbatim}
scurve_nldr_obj <- gen_scaled_data(nldr_data = scurve_umap)

scurve_nldr_obj
\end{verbatim}

\begin{verbatim}
#> $scaled_nldr
#> # A tibble: 5,000 x 3
#>     emb1  emb2    ID
#>    <dbl> <dbl> <int>
#>  1 0.707 0.839     1
#>  2 0.231 0.401     2
#>  3 0.232 0.215     3
#>  4 0.790 0.564     4
#>  5 0.761 0.551     5
#>  6 0.445 0.721     6
#>  7 0.900 0.137     7
#>  8 0.247 0.392     8
#>  9 0.325 0.542     9
#> 10 0.278 0.231    10
#> # i 4,990 more rows
#> 
#> $lim1
#> [1] -14.42166  13.32655
#> 
#> $lim2
#> [1] -12.43687  12.32455
\end{verbatim}

@@

\begin{verbatim}
calc_bins_y(
  nldr_obj = scurve_nldr_obj, 
  bin1 = 4, 
  q = 0.1)
\end{verbatim}

\begin{verbatim}
#> $bin2
#> [1] 5
#> 
#> $a1
#> [1] 0.3122301
#> 
#> $a2
#> [1] 0.2703992
\end{verbatim}

@@

\begin{verbatim}
hex_binning(
  nldr_obj = scurve_nldr_obj, 
  bin1 = 4, 
  q = 0.1)
\end{verbatim}

\begin{verbatim}
#> $a1
#> [1] 0.3122301
#> 
#> $a2
#> [1] 0.2703992
#> 
#> $bins
#> [1] 4 5
#> 
#> $start_point
#> [1] -0.10000000 -0.08923607
#> 
#> $centroids
#> # A tibble: 20 x 3
#>    hexID     c_x     c_y
#>    <int>   <dbl>   <dbl>
#>  1     1 -0.1    -0.0892
#>  2     2  0.212  -0.0892
#>  3     3  0.524  -0.0892
#>  4     4  0.837  -0.0892
#>  5     5  0.0561  0.181 
#>  6     6  0.368   0.181 
#>  7     7  0.681   0.181 
#>  8     8  0.993   0.181 
#>  9     9 -0.1     0.452 
#> 10    10  0.212   0.452 
#> 11    11  0.524   0.452 
#> 12    12  0.837   0.452 
#> 13    13  0.0561  0.722 
#> 14    14  0.368   0.722 
#> 15    15  0.681   0.722 
#> 16    16  0.993   0.722 
#> 17    17 -0.1     0.992 
#> 18    18  0.212   0.992 
#> 19    19  0.524   0.992 
#> 20    20  0.837   0.992 
#> 
#> $hex_poly
#> # A tibble: 120 x 3
#>    hex_poly_id       x         y
#>          <int>   <dbl>     <dbl>
#>  1           1 -0.1     0.0910  
#>  2           1 -0.256   0.000897
#>  3           1 -0.256  -0.179   
#>  4           1 -0.1    -0.270   
#>  5           1  0.0561 -0.179   
#>  6           1  0.0561  0.000897
#>  7           2  0.212   0.0910  
#>  8           2  0.0561  0.000897
#>  9           2  0.0561 -0.179   
#> 10           2  0.212  -0.270   
#> # i 110 more rows
#> 
#> $data_hb_id
#> # A tibble: 5,000 x 4
#>     emb1  emb2    ID hexID
#>    <dbl> <dbl> <int> <int>
#>  1 0.707 0.839     1    15
#>  2 0.231 0.401     2    10
#>  3 0.232 0.215     3     6
#>  4 0.790 0.564     4    12
#>  5 0.761 0.551     5    12
#>  6 0.445 0.721     6    14
#>  7 0.900 0.137     7     8
#>  8 0.247 0.392     8    10
#>  9 0.325 0.542     9    10
#> 10 0.278 0.231    10     6
#> # i 4,990 more rows
#> 
#> $std_cts
#> # A tibble: 16 x 3
#>    hexID bin_counts std_counts
#>    <int>      <int>      <dbl>
#>  1     3        116    0.172  
#>  2     4         82    0.122  
#>  3     5        193    0.286  
#>  4     6        674    1      
#>  5     7        529    0.785  
#>  6     8        243    0.361  
#>  7     9         87    0.129  
#>  8    10        601    0.892  
#>  9    11        244    0.362  
#> 10    12        355    0.527  
#> 11    13        183    0.272  
#> 12    14        627    0.930  
#> 13    15        669    0.993  
#> 14    16        318    0.472  
#> 15    19          2    0.00297
#> 16    20         77    0.114  
#> 
#> $tot_bins
#> [1] 20
#> 
#> $non_bins
#> [1] 16
#> 
#> $pts_bins
#> # A tibble: 16 x 2
#>    hexID pts_list   
#>    <int> <list>     
#>  1     3 <int [116]>
#>  2     4 <int [82]> 
#>  3     5 <int [193]>
#>  4     6 <int [674]>
#>  5     7 <int [529]>
#>  6     8 <int [243]>
#>  7     9 <int [87]> 
#>  8    10 <int [601]>
#>  9    11 <int [244]>
#> 10    12 <int [355]>
#> 11    13 <int [183]>
#> 12    14 <int [627]>
#> 13    15 <int [669]>
#> 14    16 <int [318]>
#> 15    19 <int [2]>  
#> 16    20 <int [77]> 
#> 
#> attr(,"class")
#> [1] "hex_bin_obj"
\end{verbatim}

@@

\begin{verbatim}
find_low_dens_hex(
  centroids_data, 
  bin1 = 15, 
  benchmark_mean_dens = 0.05
)
\end{verbatim}

@@

The final step involves lifting the fitted \(2\text{-}D\) model into \(p\text{-}D\) by computing the \(p\text{-}D\) mean of data points within each hexgaonal bin to represent bin centroids. This transformation is performed using the \texttt{avg\_highd\_data()} function, which takes \(p\text{-}D\) data (\texttt{highd\_data}) and embedding data with their corresponding hexagonal bin IDs as inputs (\texttt{scaled\_nldr\_hexid}).

\begin{verbatim}
df_bin <- avg_highd_data(
  highd_data = scurve, 
  scaled_nldr_hexid
)
\end{verbatim}

@@

\begin{verbatim}
predict_emb(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
\end{verbatim}

@@

\begin{verbatim}
glance(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
\end{verbatim}

@@

\begin{verbatim}
augment(
  highd_data = scurve, 
  model_2d = df_bin_centroids, 
  model_highd = df_bin
  )
\end{verbatim}

@@

\begin{verbatim}
vis_mesh(
  distance_edges = distance_df, 
  benchmark_value = 0.75, 
  tr_coord_df = tr_from_to_df, 
  distance_col = "distance"
  )
\end{verbatim}

@@

\begin{verbatim}
df_exe <- comb_data_mode(
  highd_data = scurve, 
  model_highd = scurve_model_obj$model_highd, 
  model_2d = scurve_model_obj$model_2d
  )
\end{verbatim}

\begin{verbatim}
show_langevitour(
  point_df, 
  edge_df
  )
\end{verbatim}
